### 2017-05-09 examples/two-diml-minimal.Rmd
start of doc for doing brownian motion

---
title: "Brownian motion on a 2D minimal system"
author: "Peter Ralph and Josh Shiffman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
```

# Model

Consider the dynamical system given by
$$\begin{aligned}
      \dot{x}(t) &= Ax(t) + Bu(t) \\
          y(t) &= Cx(t) 
\end{aligned}$$
where
$$\begin{aligned}
    B^{T} = \begin{bmatrix} 1 & 1 \end{bmatrix} \qquad C = \begin{bmatrix} 1 & 0 \end{bmatrix} .
\end{aligned}$$
and $A=A(\tau)$ depends on a free parameter,
$$\begin{aligned}
    A_0 &= \begin{bmatrix} \lambda_{1} & \lambda_{2} - \gamma \\ 0 & \lambda_{2} \end{bmatrix} \\
    V_\tau &= \begin{bmatrix} 1 & 0 \\ \tau & 1-\tau \end{bmatrix} \\
    A_\tau &= V_\tau A_0 V^{-1}_\tau
\end{aligned}$$
These systems are input-output equivalent (although may have different noise properties).

# Brownian motion

Now suppose that $A$ is changing due to mutation, 
and that it takes the value $A(t)$ after $t$ units of evolutionary time.
We know that $A(t)$ must equal $A_\tau$ for some $\tau$, 
and hence we can define $A(t) = A_{W(t)}$, 
where $W(t)$ is a random process on the free parameter;
but what are the appropriate dynamics for $W$?
Suppose that genetic drift acts uniformly on the entries of $A$, 
so that ...

# Toy example

As a toy example, suppose that we want to describe a process through $\R^2$
on the set
$$\begin{aligned}
    x^2 + y^2 = 1,
\end{aligned}$$
i.e., the unit circle.
This is parameterized by
$$\begin{aligned}
    x(\theta) &= \cos(\theta) \\
    y(\theta) &= \sin(\theta) .
\end{aligned}$$

The dynamics we want are limits of the following discrete dynamics.
Let $\epsilon$ denote the step size,
and suppose that at rate $\eta$ steps of size $\epsilon$ 
in a uniformly random direction $\phi$ are proposed;
but the probability that such a step is accepted is Kimura's expression:
$$\begin{aligned}
    q(x',y';x,y) = \frac{1 - e^{-2 (f(x',y')-f(x,y))}}{1 - e^{-2 N_e (f(x',y')-f(x,y))}}
\end{aligned}$$
where $(x',y')$ is the new (proposed) location, and
$$\begin{aligned}
    f(x,y) = (\sqrt{x^2+y^2}-1)^2
\end{aligned}$$
is the fitness function.
Note that $q(x,y;x,y) = 1$ 
and that 
$$\begin{aligned}
    \partial_u q(u,v;x,y) 
    &= 
        ( \partial_u f(u,v) ) 
        \left( \partial_z \frac{1 - e^{-2 (z-f(x,y))}}{1 - e^{-2 N_e (z-f(x,y))}} \vert_{z=f(x',y')} \right) \\
    &= 
        \left( \frac{2 x (\sqrt{x^2+y^2}-1)}{\sqrt{x^2+y^2}} \right) 
        \left( \frac{ 2 e^{-2 (f(x',y')-f(x,y))} (1 - e^{-2 N_e (f(x',y')-f(x,y))}) - 2 N_e e^{-2 (f(x',y')-f(x,y))} }{(1 - e^{-2 N_e (f(x',y')-f(x,y))})^2} \right)
\end{aligned}$$
and so that
$$\begin{aligned}
    \partial_u q(u,v;x,y) \vert_{u=x, v=y}
    &= 
        \left( \frac{2 x (\sqrt{x^2+y^2}-1)}{\sqrt{x^2+y^2}} \right) 
        \left( \frac{ - 2 N_e  }{(1 - e^{-2 N_e (f(x',y')-f(x,y))})^2} \right)
\end{aligned}$$

This has the result of multiplying jump rates by the acceptance probabilities;
so equivalently, if the process is at $(x,y)$,
then it jumps to 
$$\begin{aligned}
    (x',y') = (x+\epsilon \cos \phi, y + \epsilon \sin \phi)
\end{aligned}$$
at rate
$$\begin{aligned}
    \eta q(x',y';x,y) 
    &\approx

\end{aligned}$$