### 2017-05-25 RG_machines.tex
added notes on network size drift from our meeting

\documentclass[11 pt]{article}
\usepackage{amsmath, amssymb, color, xcolor}
\usepackage{graphicx, wrapfig, float, caption, dsfont, bbm}
\usepackage{fullpage}
\usepackage[backref=page, hidelinks, colorlinks=true, citecolor=blue!60!black!100]{hyperref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, shapes}
\usepackage{caption, subcaption}
\usepackage{natbib} % gives us \citet: Author (year) and \citep: (Author; year)
\usepackage{authblk}

\newcommand{\plr}[1]{{\color{blue}\it #1}}
\newcommand{\jss}[1]{{\color{olive}\it #1}}
% \newcommand{\ddt}{\frac{d}{dt}}
\newcommand{\ddt}{\dot}
\newcommand{\ro}{{ro}}
\newcommand{\nro}{{\bar{r}o}}
\newcommand{\rno}{{r\bar{o}}}
\newcommand{\nrno}{{\bar{r}\bar{o}}}
\newcommand{\reachable}{\mathcal{R}}
\newcommand{\unobservable}{\bar{\mathcal{O}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\pda}{\frac{\partial}{\partial A_{ij}}}
\newcommand{\ind}{\mathds{1}}

\newcommand{\A}{\mathcal{A}}
\newcommand{\diag}{\text{diag}}
\newcommand{\1}{\mathbbm{1}}

\DeclareMathOperator{\spn}{span}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}

\begin{document}

  \begin{itemize}
    \item Do gene networks drift into uneccesarily complex configurations? If the optimal dynamics of a system can be realized by $g$ genes and/or started in a state with only $g$ genes, during evolution, will the system be composed of $> g$ genes. How many more genes than necessary? 
    \item Are there forces constraining gene network size other than a fitness cost associated with the expense of uncessary energy? For instance, will genes in a non-minimal gene network contribute less overall to the system's output, diminising the cost of their removal (on average)? 
    \item as the population explores neutral network space, how often will the population be on or near a point on the manifold where a gene becomes unccessary? Does this probability go up significantly as unneccsary dimensionality goes up?
    \item Does the unncessary complexity of some gene networks confer evolvability advantages? 
  \end{itemize}
Let, 
  \begin{align*}
    \mathcal{F}_{n} = \left\{ A : C_{n}(zI-A)^{-1} B_{n} = H(z) \right\} \subseteq \mathbb{R}^{n \times n}
  \end{align*}
  be an $n$ dimensional manifold, where $A$ is gene network (not necessarily minimal), and $H(z)$ is a description of the phenotype in the Laplace domain. In the non-minimal case, $C_{k}$ and $B_{k}$ are augmented with zeroes.
  \begin{align*}
    C_{n} &= \left[ \begin{array}{cc} C_{k}  \\ \hline 0 \end{array} \right], C_{i} \neq 0 \\
      B_{n} &= \left[ \begin{array}{ccc} B_{k} & \vert & 0 \end{array} \right], B_{i} \neq 0
  \end{align*}
Let, 
  \begin{align*}
        P_{i} : \mathcal{F}_{n} \rightarrow \Omega_{n-1}
  \end{align*}
  be a projection from an $n \times n$ network to an $(n-1) \times (n-1)$ network. Biologically, this is a gene deletion or removal from the system.

  Let $G_{n,i}$ be the set of networks with identical phenotypes following a deletion. 
  \begin{align*}
        G_{n,i} = \left\{ f \in \mathcal{F}_{n} : P_{i}f \in \mathcal{F}_{n-1} \right\} \\
        G_{n} = \cup_{i} G_{n,i} \\
        d(f) = \# \left\{ i : P_{i}f \in \mathcal{F}_{n-1} \right\} 
  \end{align*}
$d(f)$ is the number of different genes that can potentially be deleted to end up in a phenotypically identical space.

Gene duplication or recruitment is defined by, 
  \begin{align*}
        U_{i}(x) \mathcal{F}_{n-1} \rightarrow \mathcal{F}_{n} \\
        \left\{ U_{i}(x) : x \in \mathbb{R}^{n} \right\} \subseteq G_{n,i} 
  \end{align*}
  where, $x$ is in the $i$th row and $0$ in the $i$th column, except along the diagonal.

  How much of a $d_{n}$-dimensional manifold is near one of $(n-k)(d_{n-1}+n)$-dimenional sub-manifolds? 
  \begin{align*}
    (n-k)\varepsilon^{d_{n}-d_{n-1} - n}
  \end{align*}

  Alternatively, what is the hitting time of $U(x)$?

\end{document}