### 2017-06-12 examples/oscillator.Rmd
oscillator example begun; more appendix notes

---
title: "Brownian motion on an oscillating system"
author: "Peter Ralph and Josh Shiffman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(expm)
library(knitr)
fig.dim <- 5
knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
```

# Model

Consider the dynamical system given by
$$\begin{aligned}
      \dot{x}(t) &= Ax(t) + Bu(t) \\
          y(t) &= Cx(t) 
\end{aligned}$$
where
$$\begin{aligned}
    B^{T} = \begin{bmatrix} 1 & 1 \end{bmatrix}, \qquad C = \begin{bmatrix} 1 & 0 \end{bmatrix} 
\end{aligned}$$
and
$$\begin{aligned}
    A = \begin{bmatrix} 0 & 1 \\ -1 & 0 \end{bmatrix} .
\end{aligned}$$
This system has transfer function
$$\begin{aligned}
    h(t) &= C e^{tA} B \\
        &= sin(t) + cos(t) .
\end{aligned}$$
The set of equivalent $A$ are given by
$\{A(\tau): \tau \neq 1 \in \R\}$, where
$$\begin{aligned}
    V(\tau) 
    &= \begin{bmatrix} 1 & 0 \\ \tau & 1-\tau \end{bmatrix} \\
    A(\tau) 
    &= V(\tau) A V(\tau)^{-1} \\
    % &= \begin{bmatrix} 0 & 1 \\ \tau-1 & \tau \end{bmatrix} 
    %    \begin{bmatrix} 1 & 0 \\ -\tau/(1-\tau) & 1/(1-\tau) \end{bmatrix}  \\
    &= \begin{bmatrix} -\tau/(1-\tau) & 1/(1-\tau) \\ -(\tau^2 + (1-\tau)^2)/(1-\tau) & \tau/(1-\tau) \end{bmatrix} 
\end{aligned}$$
These systems are input-output equivalent (although may have different noise properties).

```{r system_setup}
A0 <- matrix(c(0,-1,1,0), nrow=2)
B <- matrix(c(1,1), ncol=1)
C <- matrix(c(1,0), nrow=1)
V <- function (tau) { matrix(c(1,tau,0,1-tau), nrow=2) }
Vinv <- function (tau) { matrix(c(1,-tau/(1-tau),0,1/(1-tau)), nrow=2) }
At <- function (tau) { V(tau) %*% A0 %*% Vinv(tau) }
A <- function (tau) { 
    matrix(c(-tau/(1-tau), -(tau^2+(1-tau)^2)/(1-tau), 
             1/(1-tau), tau/(1-tau)), nrow=2) 
}
h <- function (t,tau=0) { C %*% expm(t*A(tau)) %*% B }
```

Here is the impulse response function:
```{r impulse_response_plot}
tt <- seq(0,2*pi, length.out=100)
plot(tt, sapply(tt, h), type='l',
     xlab='t', ylab=expression(h(t)), main='transfer function')
lines(tt, sin(tt)+cos(tt), col='red', lty=2)
```

# Brownian motion

Now suppose that $A$ is changing due to mutation, 
and that it takes the value $A(t)$ after $t$ units of evolutionary time.
We know that $A(t)$ must equal $A_\tau$ for some $\tau$, 
and hence we can define $A(t) = A_{W(t)}$, 
where $W(t)$ is a random process on the free parameter;
but what are the appropriate dynamics for $W$?
Suppose that genetic drift acts uniformly on the entries of $A$, 
so that ...

# Toy example

As a toy example, suppose that we want to describe a process through $\R^2$
on the set
$$\begin{aligned}
    x^2 + y^2 = 1,
\end{aligned}$$
i.e., the unit circle.
This is parameterized by
$$\begin{aligned}
    x(\theta) &= \cos(\theta) \\
    y(\theta) &= \sin(\theta) .
\end{aligned}$$

The dynamics we want are limits of the following discrete dynamics.
Let $\epsilon$ denote the step size,
and suppose that at rate $\eta$ steps of size $\epsilon$ 
in a uniformly random direction $\phi$ are proposed;
but the probability that such a step is accepted is Kimura's expression:
$$\begin{aligned}
    q(x',y';x,y) = \frac{1 - e^{-2 (f(x',y')-f(x,y))}}{1 - e^{-2 N_e (f(x',y')-f(x,y))}}
\end{aligned}$$
where $(x',y')$ is the new (proposed) location, and
$$\begin{aligned}
    f(x,y) = (\sqrt{x^2+y^2}-1)^2
\end{aligned}$$
is the fitness function.
Note that $q(x,y;x,y) = 1$ 
and that 
$$\begin{aligned}
    \partial_u q(u,v;x,y) 
    &= 
        ( \partial_u f(u,v) ) 
        \left( \partial_z \frac{1 - e^{-2 (z-f(x,y))}}{1 - e^{-2 N_e (z-f(x,y))}} \vert_{z=f(x',y')} \right) \\
    &= 
        \left( \frac{2 x (\sqrt{x^2+y^2}-1)}{\sqrt{x^2+y^2}} \right) 
        \left( \frac{ 2 e^{-2 (f(x',y')-f(x,y))} (1 - e^{-2 N_e (f(x',y')-f(x,y))}) - 2 N_e e^{-2 (f(x',y')-f(x,y))} }{(1 - e^{-2 N_e (f(x',y')-f(x,y))})^2} \right)
\end{aligned}$$
and so that
$$\begin{aligned}
    \partial_u q(u,v;x,y) \vert_{u=x, v=y}
    &= 
        \left( \frac{2 x (\sqrt{x^2+y^2}-1)}{\sqrt{x^2+y^2}} \right) 
        \left( \frac{ - 2 N_e  }{(1 - e^{-2 N_e (f(x',y')-f(x,y))})^2} \right)
\end{aligned}$$

This has the result of multiplying jump rates by the acceptance probabilities;
so equivalently, if the process is at $(x,y)$,
then it jumps to 
$$\begin{aligned}
    (x',y') = (x+\epsilon \cos \phi, y + \epsilon \sin \phi)
\end{aligned}$$
at rate
$$\begin{aligned}
    \eta q(x',y';x,y) 
    &\approx

\end{aligned}$$