### 2017-06-13 examples/oscillator.Rmd
.

---
title: "Brownian motion on an oscillating system"
author: "Peter Ralph and Josh Shiffman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(expm)
library(knitr)
fig.dim <- 5
knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
```

# Model

Consider the dynamical system given by
$$\begin{aligned}
      \dot{x}(t) &= Ax(t) + Bu(t) \\
          y(t) &= Cx(t) 
\end{aligned}$$
where
$$\begin{aligned}
    B^{T} = \begin{bmatrix} 1 & 1 \end{bmatrix}, \qquad C = \begin{bmatrix} 1 & 0 \end{bmatrix} 
\end{aligned}$$
and
$$\begin{aligned}
    A = \begin{bmatrix} 0 & 1 \\ -1 & 0 \end{bmatrix} .
\end{aligned}$$
This system has transfer function
$$\begin{aligned}
    h(t) &= C e^{tA} B \\
        &= sin(t) + cos(t) .
\end{aligned}$$
The set of equivalent $A$ are given by
$\{A(\tau): \tau \neq 1 \in \R\}$, where
$$\begin{aligned}
    V(\tau) 
    &= \begin{bmatrix} 1 & 0 \\ \tau & 1-\tau \end{bmatrix} \\
    A(\tau) 
    &= V(\tau) A V(\tau)^{-1} \\
    % &= \begin{bmatrix} 0 & 1 \\ \tau-1 & \tau \end{bmatrix} 
    %    \begin{bmatrix} 1 & 0 \\ -\tau/(1-\tau) & 1/(1-\tau) \end{bmatrix}  \\
    &= \begin{bmatrix} -\tau/(1-\tau) & 1/(1-\tau) \\ -(\tau^2 + (1-\tau)^2)/(1-\tau) & \tau/(1-\tau) \end{bmatrix} 
\end{aligned}$$
These systems are input-output equivalent (although may have different noise properties).

```{r system_setup}
A0 <- matrix(c(0,-1,1,0), nrow=2)
B <- matrix(c(1,1), ncol=1)
C <- matrix(c(1,0), nrow=1)
V <- function (tau) { matrix(c(1,tau,0,1-tau), nrow=2) }
Vinv <- function (tau) { matrix(c(1,-tau/(1-tau),0,1/(1-tau)), nrow=2) }
At <- function (tau) { V(tau) %*% A0 %*% Vinv(tau) }
A <- function (tau) { 
    matrix(c(-tau/(1-tau), -(tau^2+(1-tau)^2)/(1-tau), 
             1/(1-tau), tau/(1-tau)), nrow=2) 
}
h <- function (t,tau=0) { C %*% expm(t*A(tau)) %*% B }
```

Here is the impulse response function:
```{r impulse_response_plot}
tt <- seq(0,2*pi, length.out=100)
plot(tt, sapply(tt, h), type='l',
     xlab='t', ylab=expression(h(t)), main='transfer function')
lines(tt, sin(tt)+cos(tt), col='red', lty=2)
```

We can define the dynamics in two ways: extrinsically or intrinsically.
Extrinsically,
we look at the dynamics of a population mean in $\R^{2\times 2}$,
thinking about $A$ as evolving.
Intrinsically, we move about only on the set of equivalent $A$,
letting $\tau$ evolve.
As the population size goes to infinity,
the extrinsic dynamics should approach the intrinsic dynamics.

# Extrinsic dynamics

Let $N$ denote the population size,
and suppose that at any time the population is spread about from the mean $A$
as a Gaussian with variance $\sigma^2$, independently in each entry of $A$.
Then we expect that the population mean takes steps of size $\sigma/\sqrt{N}$
with a restoring force towards the set of optimal $A$.