### 2017-06-14 examples/oscillator.Rmd
upswitch example

---
title: "Brownian motion on an oscillating system"
author: "Peter Ralph and Josh Shiffman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(expm)
library(knitr)
library(parallel)
fig.dim <- 5
knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
do.pdfs <- TRUE
pdf_copy <- function ( width=6, height=width*knitr::opts_current$get("fig.height")/knitr::opts_current$get("fig.width"), plot.id=NULL,
                filename) {
    if (missing(filename)) {
        file.id <- if (is.null(plot.id)) { "" } else { paste0("_",plot.id) }
        filename <- knitr::fig_path(paste(file.id,".pdf",sep=""))
    }
    cat("pdf version at:",filename)
    dev.print( file=filename, device=pdf, width=width, height=height, pointsize=10, family="sans")
}
```

# Model

Consider the dynamical system given by
$$\begin{aligned}
      \dot{x}(t) &= Ax(t) + Bu(t) \\
          y(t) &= Cx(t) 
\end{aligned}$$
where
$$\begin{aligned}
    B^{T} = \begin{bmatrix} 1 & 1 \end{bmatrix}, \qquad C = \begin{bmatrix} 1 & 0 \end{bmatrix} 
\end{aligned}$$
and
$$\begin{aligned}
    A = \begin{bmatrix} 0 & 1 \\ -1 & 0 \end{bmatrix} .
\end{aligned}$$
This system has impulse response function
$$\begin{aligned}
    h(t) &= C e^{tA} B \\
        &= sin(t) + cos(t) .
\end{aligned}$$
The set of equivalent $A$ are given by
$\{A(\tau): \tau \neq 1 \in \R\}$, where
$$\begin{aligned}
    V(\tau) 
    &= \begin{bmatrix} 1 & 0 \\ \tau & 1-\tau \end{bmatrix} \\
    A(\tau) 
    &= V(\tau) A V(\tau)^{-1} \\
    % &= \begin{bmatrix} 0 & 1 \\ \tau-1 & \tau \end{bmatrix} 
    %    \begin{bmatrix} 1 & 0 \\ -\tau/(1-\tau) & 1/(1-\tau) \end{bmatrix}  \\
    &= \begin{bmatrix} -\tau/(1-\tau) & 1/(1-\tau) \\ -(\tau^2 + (1-\tau)^2)/(1-\tau) & \tau/(1-\tau) \end{bmatrix} 
\end{aligned}$$
These systems are input-output equivalent (although may have different noise properties).

```{r system_setup}
A0 <- matrix(c(0,-1,1,0), nrow=2)
B <- matrix(c(1,1), ncol=1)
C <- matrix(c(1,0), nrow=1)
V <- function (tau) { matrix(c(1,tau,0,1-tau), nrow=2) }
Vinv <- function (tau) { matrix(c(1,-tau/(1-tau),0,1/(1-tau)), nrow=2) }
At <- function (tau) { V(tau) %*% A0 %*% Vinv(tau) }
A_fn <- function (tau) { 
    matrix(c(-tau/(1-tau), -(tau^2+(1-tau)^2)/(1-tau), 
             1/(1-tau), tau/(1-tau)), nrow=2) 
}
h <- function (t,tau=0,A=A_fn(tau)) { sapply(t, function (tt) C %*% expm(tt*A) %*% B) }
optimal_h <- function (t) { sin(t)+cos(t) }
```

Here is the impulse response function:
```{r impulse_response_plot}
tt <- seq(0,2*pi, length.out=100)
plot(tt, h(tt), type='l',
     xlab='t', ylab=expression(h(t)), main='impulse response function')
lines(tt, h(tt, tau=0.1), col='blue', lty=3)
lines(tt, optimal_h(tt), col='red', lty=2)

if (do.pdfs) { pdf_copy() }
```

We can define the dynamics in two ways: extrinsically or intrinsically.
Extrinsically,
we look at the dynamics of a population mean in $\R^{2\times 2}$,
thinking about $A$ as evolving.
Intrinsically, we move about only on the set of equivalent $A$,
letting $\tau$ evolve.
As the population size goes to infinity,
the extrinsic dynamics should approach the intrinsic dynamics.

# Extrinsic dynamics

Let $N$ denote the population size,
and suppose that at any time the population is spread about from the mean $A$
as a Gaussian with variance $\sigma^2$, independently in each entry of $A$.
Then we expect that the population mean takes steps of size $\sigma/\sqrt{N}$
with a restoring force towards the set of optimal $A$.

# Fitness of hybrids

First we need the fitness function.
This will be $\exp(-D^2/2\sigma^2)$, where $D$ is the distance to optimal:
$$\begin{aligned}
  D(A) := \int_0^\infty e^{-t/4\pi} \| h(t,A) - h_0(t) \| dt ,
\end{aligned}
where $h_0(t)=\sin(t)+\cos(t)$ is the optimal impulse response function.

```{r distance_fun}
Df <- function (A,t) {
    exp(-t/(4*pi)) * ( h(t,A=A) - optimal_h(t) )^2
}
D <- function (A, upper=10, ...) {
    f <- function (t) { Df(A,t) }
    integrate(f, lower=0, upper=upper, ...)$value
}
```

To check this works, the distance of $A(2)$ to optimal should be zero:
```{r check_D}
D(A_fn(2))
```

We also need to be able to produce F1s and F2s:
```{r offspring}
F1 <- function (tau1,tau2) { (A_fn(tau1)+A_fn(tau2))/2 }

gamete <- function (X,Y,rows) { 
    out <- X
    out[rows,] <- Y[rows,]
    return(out)
}

F2 <- function (tau1, tau2, rows1, rows2) {
    (gamete(A_fn(tau1), A_fn(tau2), rows1) + gamete(A_fn(tau1), A_fn(tau2), rows2))/2
}

random_F2 <- function (tau1, tau2) {
    rows1 <- (rbinom(2,size=1,prob=1/2)>0)
    rows2 <- (rbinom(2,size=1,prob=1/2)>0)

    F2(tau1, tau2, rows1, rows2)
}

all_F2s <- function (tau1, tau2) {
    all_rows <- as.matrix(do.call( expand.grid, list(c(TRUE,FALSE))[rep(1,4)] ))
    lapply(1:nrow(all_rows), function (k) {
                x <- all_rows[k,]
                F2(tau1, tau2, x[1:2], x[3:4])
            })
}

```

Here are the phenotypes for $\tau=2$, $\tau=2.01$, and the hybrids:
```{r phenotypes}
dtau <- .01
tt <- seq(0,6*2*pi, length.out=500)
hh <- sapply( list(A_fn(2), A_fn(2+dtau), F1(2, 2+dtau)), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main=sprintf("Parental phenotypes and F1s at dtau=%s", dtau), lty=c(1,1,2), col=1:3, xlab="time", ylab="phenotype")
legend("topright", lty=c(1,1,2), col=1:3,
        legend=c(expression(tau==2), expression(tau==2+dtau), "F1"))
# now the F2s
hh <- sapply( all_F2s(2, 2+dtau), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main="F2s", xlab="time", ylab="phenotype")

if (do.pdfs) { pdf_copy() }
```
And, the same thing for $\tau=2.1$:
```{r phenotypes2}
dtau <- .1
tt <- seq(0,6*2*pi, length.out=500)
hh <- sapply( list(A_fn(2), A_fn(2+dtau), F1(2, 2+dtau)), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main=sprintf("Parental phenotypes and F1s at dtau=%s", dtau), lty=c(1,1,2), col=1:3, xlab="time", ylab="phenotype")
legend("topright", lty=c(1,1,2), col=1:3,
        legend=c(expression(tau==2), expression(tau==2+dtau), "F1"))
# now the F2s
hh <- sapply( all_F2s(2, 2+dtau), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main="F2s", xlab="time", ylab="phenotype")

if (do.pdfs) { pdf_copy() }
```

These F2s have distances of $D=$:
```{r F2_dists}
sort(round(sapply(all_F2s(2, 2+dtau), function (x) D(A=x, upper=100, subdivisions=500)),2))
```

Now let's look at $D$ as a function of $d\tau$:
```{r d_fn}

dvec <- seq(0, .02, length.out=20)
D_F1 <- sapply(dvec, function (dtau) {
            D(A=F1(tau1=2, tau2=2+dtau))
        } )

D_F2 <- do.call(cbind, mclapply(dvec, function (dtau) {
            F2s <- all_F2s(2, 2+dtau)
            sapply(F2s, D, upper=100)
        }, mc.cores=detectCores() ) )
```

```{r plot_d_fn}
layout(t(1:2))
plot( dvec, D_F1, type='l', xlab="d tau", ylab="phenotype distance (D)", main="distance of F1s")
plot( dvec, D_F1^(1/4), type='l', main="one-fourth power", xlab="d tau", ylab="D^(1/4)")

plot( dvec, D_F1, type='l', ylim=range(D_F2, D_F1),
        xlab="dtau", ylab='phenotype distance (D)',
        main="all hybrids")
lines(dvec, colMeans(D_F2), lty=2)
matlines(dvec, t(D_F2), lty=3)
legend("topleft", lty=c(1,2,3), legend=c("F1", "mean F2", "F2s"))

plot( dvec, sqrt(D_F1), type='l', ylim=sqrt(range(D_F2, D_F1)),
        xlab="dtau", ylab='phenotype distance (D)',
        main="sqrt distance")
lines(dvec, sqrt(colMeans(D_F2)), lty=2)
matlines(dvec, sqrt(t(D_F2)), lty=3)

if (do.pdfs) { pdf_copy() }
```