### 2017-12-05 kalman_decomposition.tex
edited through

\section{Kalman Decomposition} \label{apx:kalman}
\begin{definition}[Phenotypic equivalence of systems]
    Let $(\kappa(t),\phi(t))$ and $(\bar \kappa(t),\bar \phi(t))$ be the solutions to \eqref{eqn:lti_system}
    with coefficient matrices $(A,B,C)$ and $(\bar A,\bar B,\bar C)$ respectively,
    and both $\kappa(0)$ and $\bar \kappa(0)$ are zero. 
    The systems defined by $(A,B,C)$ and $(\bar A,\bar B,\bar C)$ are
    \textbf{phenotypically equivalent} 
    if
    \begin{align*}
        \phi(t) = \bar \phi(t) \qquad \text{for all} \; t \ge 0.
    \end{align*}
    Equivalently, this occurs if and only if
    \begin{align*}
        h(t) = \bar h(t)  \qquad \text{for all} \; t \ge 0,
    \end{align*}
    where $h$ and $\bar h$ are the impulse responses of the two systems.
\end{definition}

One way to find other systems equivalent to a given one
is by change of coordinates (``algebraic equivalence''):
if $V$ is an invertible matrix, then the systems $(A,B,C)$ and $(VAV^{-1},VB,CV^{-1})$
have the same dynamics because their transfer functions are equal:
\begin{align*}
    CV^{-1}( zI - VAV^{-1})^{-1}VB
    =
    CV^{-1}V( zI - A)^{-1}V^{-1}VB
    =
    C( zI - A)^{-1}B .
\end{align*}
However, the converse is not necessarily true: 
systems can have identical transfer functions without being changes of coordinates of each other.
In fact, systems with identical transfer functions can involve interactions between different
numbers of molecular species.

The set of all systems phenotypically equivalent to a given system $(A,B,C)$ 
is elegantly described using the Kalman decomposition,
which also clarifies the system dynamics? tells us a lot about how it works? \plr{or something}
To motivate this, first note that the input $u(t)$ only directly pushes the system
in directions lying in the span of the columns of $B$.
As a result, different combinations of input can 
move the system in any direction that lies in the \emph{reachable subspace},
which we denote by $\reachable$,
and is defined to be the closure of $\spn(B)$ under applying $A$
(or equivalently, the span of $B, AB, A^2B, \ldots A^{n-1}B$).
Analogously to this, we define
the \emph{observable subspace}, $\mathcal{O}$,
to be the closure of $\spn(C^T)$ under applying $A$.
(Or: $\unobservable$ is the largest $A$-invariant subspace
contained in the null space of $C$;
and $\reachable$ is the largest $A$-invariant subspace contained in the image of $B$.)

If we define
\begin{enumerate}
    \item The columns of $P_\rno$ are an orthonormal basis for $\reachable \cap \unobservable$.
    \item The columns of $P_\ro$ are an orthonormal basis of
        the complement of $\reachable \cap \unobservable$ in $\reachable$.
    \item The columns of $P_\nro$ are an orthonormal basis of
        the complement of $\reachable \cap \unobservable$ in $\unobservable$.
    \item The columns of $P_\nrno$ are an orthonormal basis of
        the remainder of $\R^n$.
\end{enumerate}
If we then define
\begin{align*}
    P &= 
    \left[ \begin{array}{c|c|c|c}
        P_\rno & P_\ro & P_\nro & P_\nrno
    \end{array} \right] ,
\end{align*}
then
\begin{align*}
    P^T P
    &=
    \left[ \begin{array}{c|c|c|c}
        I & 0 & 0 & 0 \\
        \hline
        0 & I & U & 0 \\
        \hline
        0 & V & I & 0 \\
        \hline
        0 & 0 & 0 & I 
    \end{array} \right] .
\end{align*}
\plr{Check this.  Can we get $U=V=0$?}

The following theorem can be found in SOME REFERENCE.

\begin{theorem}[Kalman decomposition] \label{thm:kalman}
        For any system $(A,B,C)$ with corresponding Kalman basis matrix $P$,
        the transformed system $(PAP^{-1},PB,CP^{-1})$  has the following form:
        \begin{align*}
            \widehat A = PAP^{-1}
            &=
            \left[ \begin{array}{cccc}
                A_{\rno} & A_{\rno,\ro} & A_{\rno,\nrno} & A_{\rno,\nro} \\
                0 & A_{\ro} & 0 & A_{\ro,\nro} \\
                0 & 0 & A_{\nrno} & A_{\nrno,\nro} \\
                0 & 0 & 0 & A_{\nro}
            \end{array} \right] ,
        \end{align*}
        and
        \begin{align*}
            \widehat B = PB
            &=
            \left[ \begin{array}{cccc}
                B_{\rno} \\
                B_{\ro} \\
                0 \\
                0 
            \end{array} \right] ,
        \end{align*}
        and
        \begin{align*}
            \widehat C = CP^{-1}
            &=
            \left[ \begin{array}{cccc}
                0 & C_{\ro} & C_{\nrno} & 0 
            \end{array} \right] .
        \end{align*}
        The transfer function of both systems is given by
        \begin{align*}
            H(z) = C_{\ro} ( zI - A_{\ro} )^{-1} B_{\ro} .
        \end{align*}
\end{theorem}

In the latter case, we say that the system is \emph{minimal} 
-- there is no equivalent system with a smaller number of species.
Note that this says that any two equivalent minimal systems
are changes of basis of each other.

Since any system can be put into this form,
and once in this form, its transfer function is determined only by 
$C_{\ro}$, $A_{\ro}$, and $B_\ro$,
therefore, the set of all equivalent systems are parameterized by
the dimension $n$,
the choice of basis ($P$),
the remaining submatrices in $\widehat A$, $\widehat B$, and $\widehat C$
(which are unconstrained),
and a invertible transformation of $\spn(P_{\ro})$, which we call $T_\ro$.

\begin{theorem}[Parameterization of equivalent systems]
    Let $(A,B,C)$ be a minimal system.
    \begin{itemize}
        \item[(a)]
            Every equivalent system is of the form given in Theorem \ref{thm:kalman},
            i.e., can be specified by choosing a dimension, $n$;
            submatrices in $\widehat A$, $\widehat B$, and $\widehat C$ 
            except for $A_\ro=A$, $B_\ro=B$, and $C_\ro=C$;
            and choosing an invertible matrix $P$.

        \item[(b)]
            \plr{conjecture:}
            The parameterization is unique
            if $P$ is furthermore chosen so that 
            each $P_x$ other than $P_\ro$ is a projection matrix,
            and that 
            \begin{align*}
                0
                =
                P_x^T P_y
            \end{align*}
            for all $(x,y)$ except $(\ro,\nrno)$.

    \end{itemize} 
\end{theorem}

\plr{Another way of saying it: pick the $\reachable$ and $\unobservable$ subspaces,
that must intersect in something of the minimal dimension;
then let $P$ be the appropriate basis?}

In some situations we may be interested in only ``network rewiring'',
where $A$ changes while $B$ and $C$ do not.
For instance, 
if all non-regulatory functions of each molecule are strongly constrained,
then $C$ cannot change.
Likewise, if responses of each molecule to the external inputs are not changed by evolution,
then $B$ does not change.

\subsection{Neutral directions from the Kalman decomposition}

The Kalman decomposition above says that any system $(A,B,C)$ can be decomposed into
$(P, \hat A, \hat B, \hat C)$ so that
$$\begin{aligned}
    A &= P^{-1} \hat A P  \\
    B &= P^{-1} \hat B  \\
    C &= \hat C P ,
\end{aligned}$$
and we know precisely how we can change these to preserve the transfer function:
\begin{enumerate}
    \item $P \to P + \epsilon Q$ as long as the result is still invertible,
    \item $\hat A \to A + \epsilon X$ as long as $X$ is zero in the correct places,
    \item $\hat B \to B + \epsilon Y$ as long as $Y$ is zero in the correct places,
    \item $\hat C \to C + \epsilon Y$ as long as $Z$ is zero in the correct places.
\end{enumerate}
By taking $\epsilon \to 0$, these tell us the local directions we can move a system $(A,B,C)$ in.
All statements below are up to first order in $\epsilon$,
omitting terms of order $\epsilon^2$.

First, since $(P + \epsilon Q)^{-1} = P^{-1} + \epsilon P^{-1} Q P^{-1}$,
modifying $P \to P + \epsilon Q$ changes
$$\begin{aligned}
    A 
        &\to A + \epsilon P^{-1} \hat A Q - \epsilon P^{-1} Q P^{-1} \hat A P \\
        &= A + \epsilon \left(A P^{-1} Q - P^{-1} Q A\right) , \\
    B
        &\to B - \epsilon P^{-1} Q B \\
    C
        &\to C + \epsilon C P^{-1} Q .
\end{aligned}$$
Since $P$ is invertible and $Q$ can be anything (if $\epsilon$ is small enough),
this allows changes in the direction of an arbitrary $W$:
$$\begin{aligned}
    A 
        &= A + \epsilon \left(A W - W A\right) , \\
    B
        &\to B - \epsilon W B \\
    C
        &\to C + \epsilon C W .
\end{aligned}$$

Then, $\hat A \to A + \epsilon X$  does
$$\begin{aligned}
    A \to A + \epsilon P^{-1} X P 
\end{aligned}$$
and $\hat B \to B + \epsilon Y$ does
$$\begin{aligned}
    B \to B + \epsilon P^{-1} Y
\end{aligned}$$
and $\hat C \to C + \epsilon Z$ does
$$\begin{aligned}
    C \to C + \epsilon Z P .
\end{aligned}$$
These degrees of freedom look like they depend on $P$, 
which is not unique,
but for any two choices of $P$ there are corresponding choices of $X$
that give the same actual change in $A$ (and likewise for $Y$ and $Z$).

Therefore, this gives us an upper bound on the number of degrees of freedom,
in terms of the dimensions of the blocks in the Kalman decomposition ($n_\ro$ etc)
and the dimensions of $B$ and $C$ ($n_B$ and $n_C$ respectively):
namely, for $W$, $X$, $Y$, and $Z$ respectively:
$$\begin{aligned}
    n^2 
    + (n_{\rno} + n_\ro n_\nro + n_\nrno(n_\nrno + n_\nro) + n_\nro^2)
    + n_B n_\rno
    + n_C n_\nrno .
\end{aligned}$$
However, some of these may be redundant.
For instance, changing $P$ in the direction of 
a $Q$ that satisfies both $A P^{-1} Q = P^{-1} Q A$ and $C P^{-1} Q = 0$
is equivalent to changing $B$ by $Y = QB$.

\section{Meiotic recombination in linear systems}\label{apx:recombination}

  Recombination is performed by taking two analogous system components from $\Sys$ and $\Sys'$ and randomly swapping rows or columns.
  \emph{E.g.} gamete systems $(A'', B'', C'')$, produced by the diploid $\{\Sys, \Sys'\}$, are formed by recombining (randomly swapping rows or columns) between two, possibly distinct, systems $\Sys = (A,B,C)$ and $\Sys' = (A', B', C')$ such that,
  \begin{align*}
    \Sys'' = \left( \begin{array}{ll}
    A'' &= MA + (I-M)A' , \\
    B'' &= MB + (I-M)B' , \\
    C'' &= CM + C'(I-M)
    \end{array} \right)
  \end{align*}
  where $M$ is a diagonal matrix where each diagonal element is a Bernoulli random variable ($M_{ii} = 0$ or $1$ with equal probability, and $M_{ij}=0$ if $i \neq j$). If systems are different dimensions the smaller system elements can be augmented with $0$s (\emph{e.g.} $\left[ \begin{smallmatrix} A & 0 \\ 0 & 0 \end{smallmatrix} \right], \left[ \begin{smallmatrix} B \\ 0 \end{smallmatrix} \right], \left[ \begin{smallmatrix} C & 0 \end{smallmatrix}\right]$).