### 2020-06-26 examples/sd_fig5.Rmd
code for fig5 plots

---
title: "sd_fig5"
author: "Josh Schiffman"
date: "6/26/2020"
output: html_document
---

```{r, cache=TRUE}
library(expm)
# System drift Figure 5 plot

A0 <- matrix(c(0,-1,1,0), nrow=2)
B <- matrix(c(1,1), ncol=1)
C <- matrix(c(1,0), nrow=1)

A_fn <- function (tau) { 
    matrix(c(tau/(tau+1), -(2*tau*(tau+1)+1)/(tau+1), 
             1/(tau+1), -tau/(tau+1)), nrow=2) 
}

tt <- seq(0,6*pi, length.out=500)

h <- function (t,tau=0,A=A_fn(tau)) { sapply(t, function (tt) C %*% expm(tt*A) %*% B) }

F1 <- function (tau1,tau2) { (A_fn(tau1)+A_fn(tau2))/2 }

get_F2s <- function(tau1, tau2) 
    {
        vc <- function(tau1, tau2, x, y) 
            {
                out <- c(A_fn(tau1)[x,y], A_fn(tau2)[x,y])
                return(out)
            }
        mlist <- list()
        count <- 1
        for (i in 1:2) 
            {
                for (j in 1:2) 
                    {
                        for (k in 1:2) 
                            {
                                for (l in 1:2)
                                    {
                                        mlist[[count]] <- matrix(c(vc(tau1, tau2, 1, 1)[i], vc(tau1, tau2, 2, 1)[j], vc(tau1, tau2, 1, 2)[k], vc(tau1, tau2, 2, 2)[l]),2,2)
count <- count + 1
                                    }
                            }
                }
            }
        f2list <- list()
        count2 <- 1
        for(i in 1:length(mlist)) 
            {
                for (j in 1:length(mlist))
                    {
                        f2list[[count2]] <- (mlist[[i]] + mlist[[j]])/2
                        count2 <- count2 + 1
                    }
            }
    return(unique(f2list))
}

hhf0 <- sapply( list(A_fn(0), A_fn(0)), function (A) { sapply(tt, h, A=A) } )

hhf1_1 <- sapply( list(A_fn(0), A_fn(0.01), F1(0, 0.01)), function (A) { sapply(tt, h, A=A) } )
hhf1_2 <- sapply( list(A_fn(0), A_fn(0.1), F1(0, 0.1)), function (A) { sapply(tt, h, A=A) } )
hhf1_3 <- sapply( list(A_fn(0), A_fn(0.5), F1(0, 0.5)), function (A) { sapply(tt, h, A=A) } )

hhf2_1 <- sapply( get_F2s(0, 0.01), function (A) { sapply(tt, h, A=A) } )
hhf2_2 <- sapply( get_F2s(0, 0.1), function (A) { sapply(tt, h, A=A) } )
hhf2_3 <- sapply( get_F2s(0, 0.5), function (A) { sapply(tt, h, A=A) } )

```

```{r, fig.align="center", fig.width=14, fig.height=10}

par(mfrow=c(3,2))
matplot(tt, hhf1_1, type='l', xlab="time", ylab="phenotype h(t)", lwd=2, ylim=c(-6,6), col=c("blue", "black"), cex.lab=1.5)
matplot(tt, hhf2_1, type='l', xlab="time", ylab="phenotype h(t)", lty=1, lwd=2, ylim=c(-6,6), cex.lab=1.5)
matlines(tt, hhf0, type='l', xlab="time", ylab="phenotype h(t)", lty=1, ylim=c(-6,6), cex.lab=1.5, col="black", lwd=4)
matplot(tt, hhf1_2, type='l', xlab="time", ylab="phenotype h(t)", lwd=2, ylim=c(-6,6), col=c("blue", "black"), cex.lab=1.5)
matplot(tt, hhf2_2, type='l', xlab="time", ylab="phenotype h(t)", lty=1, lwd=2, ylim=c(-6,6), cex.lab=1.5)
matlines(tt, hhf0, type='l', xlab="time", ylab="phenotype h(t)", lty=1, ylim=c(-6,6), cex.lab=1.5, col="black", lwd=4)
matplot(tt, hhf1_3, type='l', xlab="time", ylab="phenotype h(t)", lwd=2, ylim=c(-6,6), col=c("blue", "black"), cex.lab=1.5)
matplot(tt, hhf2_3, type='l', xlab="time", ylab="phenotype h(t)", lty=1, lwd=2, ylim=c(-6,6), cex.lab=1.5)
matlines(tt, hhf0, type='l', xlab="time", ylab="phenotype h(t)", lty=1, ylim=c(-6,6), cex.lab=1.5, col="black", lwd=4)

```